What Determines House Prices in Ames, Iowa

### 2022-07-31 Final Draft/paper.tex
proofread

	\begin{figure}
		\includegraphics[width=\textwidth]{skewplots}
		\caption{Histograms for the features. Lot Area, Lot Frontage, Total Porch Square Footage, Total Square Footage, Wood Deck Square Footage, and Years Old. Their skew values are 1.54, 12.59, 1.55, 0.65, 2.01, and 0.61, respectively.}
		\label{fig:skew}
	\end{figure}

@@

	Looking at the categorical we plot the averages against each other to get a visual idea of whats going on. Visually looking at the plots, we see that Neighborhood and Overall Quality look to have statistically significant different means.

	\begin{figure}
		\includegraphics[width=\textwidth]{catboxes}
		\caption{Box Plots of all categorical variables. Neighborhood, Overall Qual, and Overall Condition have noticably different means.}
		\label{fig:cats}
	\end{figure}

@@

	\begin{table}[]
		\centering
		\begin{tabular}{llll}
		\hline
		\multicolumn{1}{|l|}{Feature} & \multicolumn{1}{l|}{F-Stat} & \multicolumn{1}{l|}{P-value} & \multicolumn{1}{l|}{Significant?} \\ \hline
		Overall Quality               & 374.11                      & 0.00                         & True                              \\
		Neighborhood                  & 74.02                       & 1.63e-230                    & True                              \\
		Overall Condition             & 28.84                       & 1.87e-41                     & True                              \\
		Month Sold                    & 0.98                        & 0.458                        & False                             \\
		Year Sold                     & 0.32                        & 0.865                        & False                            
		\end{tabular}
		\end{table}

@@

	\section{Modelling}

@@

	As stated in previous the "Skewness of Variables" section, we first log transform the features LotFrontage, Lot Area, Wood Deck Square Footage, Total Square Footage, Total Porch Square Footage, and Years Old. In addidtion to those features, we log transform the target varaible, Sales Price.

@@

	\begin{table}[]
		\centering
		\begin{tabular}{ll}
		\hline
		\multicolumn{1}{|l|}{feature} & \multicolumn{1}{l|}{VIF} \\ \hline
		LotFrontage                   & 2.29459                  \\
		LotArea                       & 2.854278                 \\
		WoodDeckSF                    & 1.244522                 \\
		GarageArea                    & 1.997873                 \\
		TotalSF                       & 3.169309                 \\
		TotalPorchSF                  & 1.272145                 \\
		TotalBaths                    & 2.113038                 \\
		YearsOld                      & 7.157472                
		\end{tabular}
		\end{table}

@@

	\begin{table}[]

@@

	From the above table, we can see that in general, the model is pretty effective. With and Adjusted R-Squared of 0.909, we are explaining almost 91 percent of the variance. Our plot of residuals may be slightly skewed to the positive residuals, but nothing of too much concern. The p-value of the model was too low for the software to discern, so we are left with a p-value that is virtually 0.

	\begin{figure}
		\includegraphics[width=\textwidth]{mlrplot}
		\caption{Residual Plots for Multiple Linear Regression}
		\label{fig:skew}
	\end{figure}

@@

	\begin{figure}
		\includegraphics[width=\textwidth]{ridgeform}
		\caption{Formula for Ridge Regression. Is the same as OLS Regression, but a shrinkage penalty is applied.}
		\label{fig:skew}
	\end{figure}

	Inorder to choosed the value for our tuning parameter, we decided to run a 5-Fold Cross validation. After performing said cross validation, it was determined that the best alpha level for the Ridge Regression model was 0.7. After performing ridge regression, here was our output, as well as the features who's coefficients were given the most weight:

@@

	After performming Ridge Regression, we were able to increase our Adjusted R-Squred only slightly, to 0.9103. However, what we gain more insight from is the variable's that it weight the most heavily. Those features were almost all measure of Quality and Condition. The rankings at the top and bottom of the spectrum, 1, 2, 9, and 10, were given the most weight.

@@

	\begin{figure}
		\includegraphics[width=\textwidth]{lassoform}
		\caption{Formula for LASSO Regression. Is the same as OLS Regression, but a shrinkage penalty is applied.}
		\label{fig:skew}
	\end{figure}

	Inorder to choosed the value for our tuning parameter, we decided to run a 5-Fold Cross validation. After performing said cross validation, it was determined that the best alpha level for the LASSO Regression model was 5e-05. After performing LASSO Regression, here was our output, as well as the features who's coefficients were given the most weight:

@@

	The LASSO Regression model dropped 3 of the 52 features, those being: Lot Frontage, Neighborhood SWISU, and Overall Condition 5. This comes as no surprise as again the most highly weighted features were the Overall Quality and Overall Condition scores at each end of the rating scale. Additionaly, recall that the MLR found that Lot Frontage was not significant, so that could explain why LASSO reduced it to 0.

@@

	From our three models, we can conclude that the Ridge Regression performed the best of the 3. See the table below for the Adjusted-R Squared values for each model. Additionally, we see that the least important variable that we passed into each model was Lot Frontage.
	From the Ridge and LASSO regressions, we can see that Overall Condition, and Overall Quality are the most important factors in determining the price of a home. What's interesting, is that it is only the ratings for Condition and Quality at the top and bottom ends of the spectrum that are given any weight. This must mean that if you are willing to rate a house very poorly for Quality and Condition, it's price will almost assuredly be less. If you are blown away by the home and willing to rate it a 9 or a 10, it will assuredly cost more. However, the more "average" ratings, let's say 3-7, do not influence the price of a home as severly. It is also important to note that the old adage of "Location, Location, Location" does seem to be true as Loactions were the coefficients with the third highest weighting in both the Ridge and LASSO regressions.

@@

	We set out to prove whether Size, Condition, Quality, Location, and Age were actually significant factors in the price of a home. It is clear that these variables are indeed the main factors in determining home prices. Quality, then Condition, and then Location are the most influential. Because we reduced the original dataset, it might be beneficial to redo this study with more features. While this make the models less interpretable, there may be other important factors that influence home prices that we have not thought of. In addition, expanding this study to include more homes in more locations. What affects home prices could be largely regional. For all we know, it could change from town to town, nevermind state to state or country to country. Nonetheless, it would appear that Quality, Condition, Size, and Age are the driving factors in home price.

>>>>

	\begin{figure}[H]
		\includegraphics[width=\textwidth]{skewplots}
		\caption{Histograms for the features. Lot Area, Lot Frontage, Total Porch Square Footage, Total Square Footage, Wood Deck Square Footage, and Years Old. Their skew values are 1.54, 12.59, 1.55, 0.65, 2.01, and 0.61, respectively.}
		\label{fig:skew}
	\end{figure}

@@

	Looking at the categorical variables we plot the averages against each other to get a visual idea of whats going on. Visually looking at the plots, we see that Neighborhood and Overall Quality look to have statistically significant different means.

	\begin{figure}[H]
		\includegraphics[width=\textwidth]{catboxes}
		\caption{Box Plots of all categorical variables. Neighborhood, Overall Qual, and Overall Condition have noticably different means.}
		\label{fig:cats}
	\end{figure}

@@

	\begin{table}[H]
		\centering
		\begin{tabular}{llll}
		\hline
		\multicolumn{1}{|l|}{Feature} & \multicolumn{1}{l|}{F-Stat} & \multicolumn{1}{l|}{P-value} & \multicolumn{1}{l|}{Significant?} \\ \hline
		Overall Quality               & 374.11                      & 0.00                         & True                              \\
		Neighborhood                  & 74.02                       & 1.63e-230                    & True                              \\
		Overall Condition             & 28.84                       & 1.87e-41                     & True                              \\
		Month Sold                    & 0.98                        & 0.458                        & False                             \\
		Year Sold                     & 0.32                        & 0.865                        & False                            
		\end{tabular}
		\end{table}

@@

	\section{Modeling}

@@

	As stated in previous the "Skewness of Variables" section, we first log transform the features LotFrontage, Lot Area, Wood Deck Square Footage, Total Square Footage, Total Porch Square Footage, and Years Old. In addition to those features, we log transform the target variable, Sales Price.

@@

	\begin{table}[H]
		\centering
		\begin{tabular}{ll}
		\hline
		\multicolumn{1}{|l|}{feature} & \multicolumn{1}{l|}{VIF} \\ \hline
		LotFrontage                   & 2.29459                  \\
		LotArea                       & 2.854278                 \\
		WoodDeckSF                    & 1.244522                 \\
		GarageArea                    & 1.997873                 \\
		TotalSF                       & 3.169309                 \\
		TotalPorchSF                  & 1.272145                 \\
		TotalBaths                    & 2.113038                 \\
		YearsOld                      & 7.157472                
		\end{tabular}
		\end{table}

@@

	\begin{table}[H]

@@

	From the above table, we can see that in general, the model is pretty effective. With an Adjusted R-Squared of 0.909, we are explaining almost 91 percent of the variance. Our plot of residuals may be slightly skewed to the positive residuals, but nothing of too much concern. The p-value of the model was too low for the software to discern, so we are left with a p-value that is virtually 0.

	\begin{figure}[H]
		\includegraphics[width=\textwidth]{mlrplot}
		\caption{Residual Plots for Multiple Linear Regression}
		\label{fig:skew}
	\end{figure}

@@

	\begin{figure}[H]
		\includegraphics[width=\textwidth]{ridgeform}
		\caption{Formula for Ridge Regression. Is the same as OLS Regression, but a shrinkage penalty is applied.}
		\label{fig:skew}
	\end{figure}

	In order to choose the value for our tuning parameter, we decided to run a 5-Fold Cross validation. After performing said cross validation, it was determined that the best alpha level for the Ridge Regression model was 0.7. After performing ridge regression, here was our output, as well as the features who's coefficients were given the most weight:

@@

	After performing Ridge Regression, we were able to increase our Adjusted R-Squared only slightly, to 0.9103. However, what we gain more insight from is the variable's that it weighs the most heavily. Those features were almost all measure of Quality and Condition. The rankings at the top and bottom of the spectrum, 1, 2, 9, and 10, were given the most weight.

@@

	\begin{figure}[H]
		\includegraphics[width=\textwidth]{lassoform}
		\caption{Formula for LASSO Regression. Is the same as OLS Regression, but a shrinkage penalty is applied.}
		\label{fig:skew}
	\end{figure}

	In order to choose the value for our tuning parameter, we decided to run a 5-Fold Cross validation. After performing said cross validation, it was determined that the best alpha level for the LASSO Regression model was 5e-05. After performing LASSO Regression, here was our output, as well as the features who's coefficients were given the most weight:

@@

	The LASSO Regression model dropped 3 of the 52 features, those being: Lot Frontage, Neighborhood SWISU, and Overall Condition 5. This comes as no surprise as again the most highly weighted features were the Overall Quality and Overall Condition scores at each end of the rating scale. Additionally, recall that the MLR found that Lot Frontage was not significant, so that could explain why LASSO reduced it to 0.

@@

	From our three models, we can conclude that the Ridge Regression performed the best of the 3. See the table below for the Adjusted-R Squared values for each model. Additionally, we see that the least important variable that we passed into each model was Lot Frontage.
	From the Ridge and LASSO regressions, we can see that Overall Condition, and Overall Quality are the most important factors in determining the price of a home. What's interesting, is that it is only the ratings for Condition and Quality at the top and bottom ends of the spectrum that are given any weight. This must mean that if you are willing to rate a house very poorly for Quality and Condition, it's price will almost assuredly be less. If you are blown away by the home and willing to rate it a 9 or a 10, it will assuredly cost more. However, the more "average" ratings, let's say 3-7, do not influence the price of a home as severly. It is also important to note that the old adage of "Location, Location, Location" does seem to be true as Locations were the coefficients with the third highest weighting in both the Ridge and LASSO regressions.

@@

	We set out to prove whether Size, Condition, Quality, Location, and Age were actually significant factors in the price of a home. It is clear that these variables are indeed the main factors in determining home prices. Quality, then Condition, and then Location are the most influential. Because we reduced the original data set, it might be beneficial to redo this study with more features. While this make the models less interpretable, there may be other important factors that influence home prices that we have not thought of. In addition, expanding this study to include more homes in more locations. What affects home prices could be largely regional. For all we know, it could change from town to town, nevermind state to state or country to country. Nonetheless, it would appear that Quality, Condition, Size, and Age are the driving factors in home price.
